Quantifying the home advantage with `hyper3` objects

### 2023-02-08 inst/home_advantage3.Rmd
typos and mathematica transcript

$$
\hat{p_1}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{p_2}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}
$$

@@

$$
\hat{p_1}=\frac{C}{C+D}\qquad
\hat{p_2}=\frac{A}{A+B}\qquad
M=\frac{AD-BC}{(A+B)(B+C)}
$$

@@

$$
\hat{p_1}=\frac{\sqrt{21}}{\sqrt{21} + \sqrt{35}}\simeq 0.47\qquad
\hat{p_2}=\frac{\sqrt{63}}{\sqrt{21} + \sqrt{35}}\simeq 0.53\qquad
\hat{\lambda}=\sqrt{\frac{AD}{BC}}=2.05.
$$

@@

>>>>

$$
\hat{p_1}=\frac{\sqrt{AD}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{p_2}=\frac{\sqrt{BC}}{\sqrt{AD}+\sqrt{BC}}\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}
$$

@@

$$
\hat{p_1}=\frac{C}{C+D}\qquad
\hat{p_2}=\frac{B}{A+B}\qquad
M=\frac{AC-BD}{(A+B)(C+D)}
$$

@@

$$
\hat{p_1}=\frac{\sqrt{21}}{\sqrt{21} + \sqrt{35}}\simeq 0.47\qquad
\hat{p_2}=\frac{\sqrt{63}}{\sqrt{21} + \sqrt{35}}\simeq 0.53\qquad
\hat{\lambda}=\sqrt{\frac{AC}{BD}}=2.05.
$$

@@

# Appendix

Here is a transcript of a mathematica session which establishes the
estimates above.  First the `hyper2` likelihood:

```
rhankin@rhrpi4:~ $ wolfram
Mathematica 12.2.0 Kernel for Linux ARM (32-bit)
Copyright 1988-2021 Wolfram Research, Inc.

In[1]:= L = Log[M+p1]*A +Log[p2]*B- Log[M+p1+p2]*(A+B+C+D) + Log[p2+M]*C + Log[p1]*D

Out[1]= D Log[p1] + A Log[M + p1] + B Log[p2] + C Log[M + p2] - (A + B + C + D) Log[M + p1 + p2]

In[2]:= LL = L /. {p2 -> 1-p1-M}

Out[2]= C Log[1 - p1] + B Log[1 - M - p1] + D Log[p1] + A Log[M + p1]

In[3]:= dp = D[LL,p1]//FullSimplify

            C      D         B          A
Out[3]=  ------- + -- + ----------- + ------
         -1 + p1   p1   -1 + M + p1   M + p1

In[4]:= dM = D[LL,M]//FullSimplify

              B          A
Out[4]= ----------- + ------
         -1 + M + p1   M + p1

In[5]:= Solve[{dp==0,dM==0},{p1,M}]

                   D            A C - B D
Out[5]=  {{p1 -> -----, M -> ---------------}}
                 C + D       (A + B) (C + D)

In[6]:= p2 = 1-p1-M /. {%} //FullSimplify

             B
Out[6]=  {{-----}}
           A + B

```

Now the `hyper3` likelihood:

```
rhankin@rhrpi4:~ $ wolfram
Mathematica 12.2.0 Kernel for Linux ARM (32-bit)
Copyright 1988-2021 Wolfram Research, Inc.

In[1]:= L = A*Log[lambda*p1] + B*Log[p2] -(A+B)*Log[lambda*p1 + p2] + C*Log[lambda*p2] + D*Log[p1] -(C+D)*Log[p1 + lambda*p2]

Out[1]= D Log[p1] + A Log[lambda p1] + B Log[p2] + C Log[lambda p2] - (A + B) Log[lambda p1 + p2] - (C + D) Log[p1 + lambda p2]

In[2]:= LL = L /. {p2 -> 1-p1}

Out[2]= B Log[1 - p1] + C Log[lambda (1 - p1)] + D Log[p1] + A Log[lambda p1] - (C + D) Log[lambda (1 - p1) + p1] -

>    (A + B) Log[1 - p1 + lambda p1]

In[3]:= dp1 = D[LL,p1] //FullSimplify

         B + C    A + D   (A + B) (-1 + lambda)    (C + D) (-1 + lambda)
Out[3]= ------- + ----- - --------------------- + -----------------------
        -1 + p1    p1     1 + (-1 + lambda) p1    lambda + p1 - lambda p1

In[4]:= dlam = D[LL,lambda]//FullSimplify

          A        C           (A + B) p1           (C + D) (-1 + p1)
Out[4]= ------ + ------ - -------------------- + -----------------------
        lambda   lambda   1 + (-1 + lambda) p1   lambda + p1 - lambda p1

In[5]:= Solve[{dp1==0,dlam==0},{p1,lambda}]//FullSimplify

[snip]

In[6]:= %[[1]]

                        1                     Sqrt[A] Sqrt[C]
Out[6]= {p1 -> -------------------, lambda -> ---------------}
                   Sqrt[B] Sqrt[C]            Sqrt[B] Sqrt[D]
               1 + ---------------
                   Sqrt[A] Sqrt[D]

```
